### 2018-01-10 project/include/tikz/distance_real_line.tikz.tex

\begin{tikzpicture}
\tikzmath{
  \a = 3;
  \b = 2;
  \i0 = 0;
  \i1 = 0*\a+\b;
  \i2 = 0.75*\a+\b;
  \i3 = 1.7*\a+\b;
  \i4 = 2.5*\a+\b;
  \i5 = 3.25*\a+\b;
  \i6 = 4*\a+\b;
  \s0 = 0;
  \s1 = \s0 + \i1;
  \s2 = \s1 + \i2;
  \s3 = \s2 + \i3;
  \s4 = \s3 + \i4;
  \s5 = \s4 + \i5;
  \s6 = \s5 + \i6;
}

% circles
\begin{scope}
  \clip (0,-1.5) rectangle (\i6+1,1.5);
  \draw[dashed,green] (0,0) circle [radius=\i1];
  \draw[dashed,green] (0,0) circle [radius=\i2];
  \draw[dashed,green] (0,0) circle [radius=\i3];
  \draw[dashed,green] (0,0) circle [radius=\i4];
  \draw[dashed,green] (0,0) circle [radius=\i5];
  \draw[dashed,green] (0,0) circle [radius=\i6];
\end{scope}

%\draw[-latex] (0,0) -- (4.5*\a+\b,0);
\draw[-] (0,0) -- (\i1*0.2,0);
\draw[dashed] (\i1*0.2,0) -- (\i1-\i1*0.2,0);
\draw[-] (\i1-\i1*0.2,0) -- (\i1,0);

\draw[-] (\i1,0) -- (\i1+\i2*0.2-\i1*0.2,0);
\draw[dashed] (\i1+\i2*0.2-\i1*0.2,0) -- (\i2-\i2*0.2+\i1*0.2,0);
\draw[-] (\i2-\i2*0.2+\i1*0.2,0) -- (\i2,0);

\draw[-,thick,red] (\i2,0) -- (\i3,0);

\draw[-] (\i3,0) -- (\i3+\i4*0.2-\i3*0.2,0);
\draw[dashed] (\i3+\i4*0.2-\i3*0.2,0) -- (\i4-\i4*0.2+\i3*0.2,0);
\draw[-] (\i4-\i4*0.2+\i3*0.2,0) -- (\i4,0);

\draw[-] (\i4,0) -- (\i5,0);

\draw[-] (\i5,0) -- (\i5+\i6*0.2-\i5*0.2,0);
\draw[dashed] (\i5+\i6*0.2-\i5*0.2,0) -- (\i6-\i6*0.2+\i5*0.2,0);
\draw[-] (\i6-\i6*0.2+\i5*0.2,0) -- (\i6,0);

\draw[-latex] (\i6,0) -- (\i6+1,0);

% 0
\draw[shift={(0,0)},color=black] (0pt,3pt) -- (0pt,-3pt);
\draw[shift={(0,0)},color=black] (0pt,0pt) -- (0pt,-3pt) node[below] 
{$0$};

% p_i_1
\draw[shift={(\i1,0)},color=black] (0pt,3pt) -- (0pt,-3pt);
\draw[shift={(\i1,0)},color=black] (0pt,0pt) -- (0pt,-3pt) node[below] 
{$d_G(p_{i_1}(u))$};

% p_j
\draw[shift={(\i2,0)},color=black] (0pt,3pt) -- (0pt,-3pt);
\draw[shift={(\i2,0)},color=black] (0pt,0pt) -- (0pt,-3pt) node[below] 
{$d_G(p_j(u))$};

% p_j_p2
\draw[shift={(\i3,0)},color=black] (0pt,3pt) -- (0pt,-3pt);
\draw[shift={(\i3,0)},color=black] (0pt,0pt) -- (0pt,-3pt) node[below] 
{$d_G(p_{j+2}(u))$};

% p_i
\draw[shift={(\i4,0)},color=black] (0pt,3pt) -- (0pt,-3pt);
\draw[shift={(\i4,0)},color=black] (0pt,0pt) -- (0pt,-3pt) node[below] 
{$d_G(p_{i-2}(u))$};

% p_i_p2
\draw[shift={(\i5,0)},color=black] (0pt,3pt) -- (0pt,-3pt);
\draw[shift={(\i5,0)},color=black] (0pt,0pt) -- (0pt,-3pt) node[below] 
{$d_G(p_{i}(u))$};

% p_i_2
\draw[shift={(\i6,0)},color=black] (0pt,3pt) -- (0pt,-3pt);
\draw[shift={(\i6,0)},color=black] (0pt,0pt) -- (0pt,-3pt) node[below] 
{$d_G(p_{i_2}(u))$};

% d_G(u,p_{1_i}(u))
\draw[decorate,decoration={brace,amplitude=10pt,raise=4pt},yshift=0pt]
(0,1.2) -- (\i1,1.2) node [black,midway,xshift=0cm,yshift=23pt] {\footnotesize
$d_G(u,p_{i_1}(u))$};

% \delta_j(u)
\draw[decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
(\i3,0.075) -- (\i2,0.075) node [black,midway,xshift=0cm,yshift=23pt] {\footnotesize
$\delta_j(u)$};

% \sum
\draw[decorate,decoration={brace,amplitude=10pt,mirror,raise=4pt},yshift=0pt]
(\i5,1.2) -- (\i1,1.2) node [black,midway,xshift=0cm,yshift=23pt] {\footnotesize
$\sum_{j' \in \mathcal{I}'}\delta_{j'}(u)$};

\draw[dashed] (0,0) -- (0,1.2);
\draw[dashed] (\i1,0) -- (\i1,1.2);
\draw[dashed] (\i5,0) -- (\i5,1.2);

\end{tikzpicture}