### 2018-01-10 project/include/tikz/lemma1.tikz.tex
asd

\begin{tikzpicture}
  \filldraw (0,0) circle (2pt) node[align=left,   below] {$u$};
  \filldraw (10,0) circle (2pt) node[align=right,  below] {$v$};
  \filldraw[fill=gray] (4,3) circle (2pt) node[align=left, above] {$p_i(u)$};
  \filldraw[fill=white] (8,3.45) circle (2pt) node[align=left, above] {$p_{i+1}(v)$};
  \coordinate (u) at (0,0) ;
  \coordinate (v) at (10,0);
  \coordinate (pu) at (4,3);
  \coordinate (pv) at (8, 3.45);
  \begin{scope}
    \clip (0,6) rectangle (7,-1);
    \draw[dashed,green] (u) circle [radius=5];
  \end{scope}
  \draw (u) -- (pu) node[pos=0.5, above, sloped] {$d_G(u,p_i(u))$};
  \draw (u) -- (v) node[pos=0.5, below, sloped] {$d_G(u,v))$};
  \draw (v) -- (pv) node[pos=0.5, below, sloped] {$d_G(v, p_{i+1}(v)$};
\end{tikzpicture}